Categories of Optics

### 2018-06-11 paper.tex
Reindent everything

\documentclass[11pt,a4paper]{article}
\usepackage{authblk}
%\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
%\usepackage{stmaryrd}
%\usepackage{tensor}
\usepackage[mathscr]{eucal}
\usepackage{url}
%\usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{minted}

@@

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[subsection]
%\newtheorem{axiom}[theorem]{Axiom}
\newtheorem*{theoremstar}{Theorem}
%\newtheorem{fact}[theorem]{Fact}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
%\newtheorem{convention}[theorem]{Convention}
%\newtheorem{construction}[theorem]{Construction}
\newtheorem{example}[theorem]{Example}
%\newtheorem{examples}[theorem]{Examples}
%\newtheorem{notation}[theorem]{Notation}
\newtheorem{remark}[theorem]{Remark}
%\newtheorem{idea}[theorem]{Idea}
%\newtheorem{question}[theorem]{Question}

@@

\newcommand{\Pastro}{\Phi}
%\newcommand{\Pastro}{\mathrm{Pastro}}
\newcommand{\Double}{\mathcal{D}}

@@

\newif\ifhideproofs
%\hideproofstrue %uncomment to hide proofs
\ifhideproofs
\usepackage{environ}
\NewEnviron{hide}{}
\let\proof\hide
\let\endproof\endhide
\fi

@@

If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for the representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

\begin{example}
For any three objects $M, A, A' \in \C$, there is the \emph{tautological} optic \[t_{M,A} : (M \otimes A, M \otimes A') \hto (A, A')\] which is given by $\langle \id_{M \otimes A}, \id_{M \otimes A'} \rangle$.
\end{example}

\begin{remark}
Elements of $\Optic_\C((S, S'), (A, A'))$ have an appealing interpretation as string diagrams with a ``hole'' missing. We draw the pair $\langle l, r \rangle$ as
\begin{center}
\input{diagrams/generic-optic.tikz}
\end{center}
reading left to right, so the left half of the diagram represents $l$ and the right half $r$. The relation imposed by the coend equates the following two diagrams:
\begin{center}
\input{diagrams/coend-relation-left.tikz} 
\qquad
\input{diagrams/coend-relation-right.tikz}
\end{center}
We will therefore omit the vertical cut point between $l$ and $r$ in subsequent diagrams; any choice yields a representative of same optic.
\end{remark}

Two optics can be composed as follows: \todo{I give possibly too much detail here.} We wish to construct a map
\begin{align*}
&\left(\int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')\right) \times \left(\int^{M \in \C} \C(R, M \otimes S) \times \C(M \otimes S', R')\right) \\ &
 \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
The product in $\Set$ preserves colimits, so in particular coends. Using this fact, and the Fubini theorem for coends \cite[Remark 1.9]{CoendCofriend}, the domain is isomorphic to
\begin{align*}
\int^{(M, N) \in \C \times \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R').
\end{align*}
So by the universal property of coends, it suffices to construct natural maps 
\begin{align*}
& \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') \\ &
 \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
for each $M, N \in \C$. For this we use the composites
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R')\\
\to \,& \C(N \otimes S, N \otimes M \otimes A) \times \C(N \otimes M \otimes A', N \otimes S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') && \text{(functoriality of $N \otimes  -$)} \\
\to \,& \C(R, N \otimes  M \otimes A) \times \C(N \otimes M \otimes A', R') && \text{(composition in $\C$)} \\
\to \,&\int^{M \in \M} \C(R, M \otimes A) \times \C(M \otimes A', R') && \text{($\copr_{N \otimes M}$)}
\end{align*}
In other words, let $\langle l', r' \rangle : (R, R') \hto (S, S')$ and $\langle l, r \rangle : (S, S') \hto (A, A')$ with $M$ the complement for $\langle l', r' \rangle$. The composite is then: $\langle l, r \rangle \circ \langle l', r' \rangle = \langle (M l)l', r'(Mr)\rangle$.

@@

\begin{proposition}
The above data form a category $\Optic_\C$.
\end{proposition}
\begin{proof}
In \cite[Section 6]{Doubles} this is proven abstractly, by exhibiting this category as the Kleisli category for a monad in the bicategory $\Prof$. Here we prove the properties directly.

We check associativity of composition equationally. As in the definition of composition, the Fubini theorem for coends allows us to choose representatives for all three optics simultaneously. Suppose we have representatives of three optics $\langle l_1, r_1\rangle : (R, R) \hto (S, S')$ and $\langle l_2, r_2 \rangle : (S, S') \hto (A, A')$ and $\langle l_3, r_3 \rangle : (A, A') \hto (B, B')$, that have complements $M$, $N$ and $P$ respectively. Then:
\begin{align*}
(\langle l_3, r_3 \rangle \circ \langle l_2, r_2 \rangle) \circ \langle l_1, r_1 \rangle 
&= \langle (N \otimes l_3)l_2, r_2(N \otimes r_3) \rangle \circ \langle l_1, r_1 \rangle \\
&= \langle (M \otimes ((N \otimes l_3)l_2))l_1, r_1(M \otimes (r_2(N \otimes r_3))) \rangle \\
&= \langle (M \otimes N \otimes l_3)(M \otimes l_2)l_1, r_1(M \otimes r_2)(M \otimes N \otimes r_3)) \rangle \\
&= \langle l_3, r_3 \rangle \circ (\langle (M \otimes l_2)l_1, r_1(M \otimes r_2) \rangle) \\
&= \langle l_3, r_3 \rangle \circ (\langle l_2, r_2 \rangle \circ \langle l_1, r_1 \rangle)
\end{align*}

@@

The remainder of this section comprises some useful observations that were not made in \cite{Doubles}.

@@

\begin{remark}
This functor is not necessarily faithful, see Remark \ref{lens-iota-not-faithful}.
\end{remark}

%\todo{This doesn't give us the structure of a `framing', we can only lift vertical arrows to horizontal arrows facing one direction.}

%\begin{theorem}
%$\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, S' \otimes T')$, the unit is $(I, I)$, and the action on morphisms induced by
%\begin{align*}
%&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
%\to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
%\to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
%\to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}
%\end{align*}
%\end{theorem}
%\begin{proof}
%Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
%\begin{align*}
%\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
%\end{align*}
%This is functorial, as suggested by the equivalence of the following string diagrams:
%\todo{todo}
%%For future convenience, note that 
%%\begin{align*}
%%\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
%%&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
%%&= \langle l \otimes T, r \otimes T' \rangle
%%\end{align*}
%%Now we check functoriality. Suppose we have:
%%\begin{align*}
%%\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
%%\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
%%\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
%%\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
%%\end{align*}
%%with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
%%\begin{align*}
%%&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
%%= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
%%&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
%%&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
%%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
%%= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
%%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
%%= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
%%= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
%%\end{align*}

@@

%The structure morphisms are all lifted from the structure morphisms in $\C$:
%\begin{align*}
%\alpha_{(R, R'), (S, S'), (T, T')} &:= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1}) \\
%\lambda_{(S, S')} &:= \iota(\lambda_{(S, S')}, \lambda_{(S, S')}^{-1}) \\
%\rho_{(S, S')} &:= (\rho_{(S, S')}, \rho_{(S, S')}^{-1}) \\
%s_{(S, S'), (T, T')} &:= \iota(s_{S, T}, s_{T', S'})
%\end{align*}

@@

%Note that because $\iota(S, S') = (S, S')$, required equations for $\iota$ to be a monoidal functor hold by definition (although we don't yet know that $\Optic_\otimes$ is monoidal). The pentagon and triangle axioms then hold in $\Optic_\otimes$, as they are the image of the same axioms in $\C \times \C^\op$ under $\iota$. The only remaining thing to verify is that these structure maps are natural in $\Optic_\otimes$. 

@@

%Let us check that $\alpha_{(R, R'), (S, S'), (T, T')}$ is natural in $(R, R')$. Suppose $f : (Q, Q') \hto (R, R')$ has representative $(l, r) : (Q \to M \otimes R, M \otimes R' \to Q')$, then:
%\begin{align*}
%&\alpha_{(R, R'), (S, S'), (T, T')} \circ (f \otimes (S, S')) \otimes (T, T') \\
%&= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1})  \circ (\langle l, r \rangle \otimes (S, S')) \otimes (T, T') \\
%&= \langle\lambda_{R \otimes (S \otimes T)}\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1} \lambda^{-1}_{(R' \otimes S') \otimes T'} \rangle  \circ \left\langle (l \otimes S) \otimes T, (r \otimes S') \otimes T' \right\rangle \\

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage{authblk}
% \usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
% \usepackage{stmaryrd}
% \usepackage{tensor}
\usepackage[mathscr]{eucal}
\usepackage{url}
% \usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{minted}

@@

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[subsection]
% \newtheorem{axiom}[theorem]{Axiom}
\newtheorem*{theoremstar}{Theorem}
% \newtheorem{fact}[theorem]{Fact}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
% \newtheorem{convention}[theorem]{Convention}
% \newtheorem{construction}[theorem]{Construction}
\newtheorem{example}[theorem]{Example}
% \newtheorem{examples}[theorem]{Examples}
% \newtheorem{notation}[theorem]{Notation}
\newtheorem{remark}[theorem]{Remark}
% \newtheorem{idea}[theorem]{Idea}
% \newtheorem{question}[theorem]{Question}

@@

\newcommand{\Pastro}{\Phi}
% \newcommand{\Pastro}{\mathrm{Pastro}}
\newcommand{\Double}{\mathcal{D}}

@@

\newif\ifhideproofs
% \hideproofstrue %uncomment to hide proofs
\ifhideproofs
\usepackage{environ}
\NewEnviron{hide}{}
\let\proof\hide
\let\endproof\endhide
\fi

@@

If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l \mid r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for the representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

\begin{example}
  For any three objects $M, A, A' \in \C$, there is the \emph{tautological} optic \[t_{M,A} : (M \otimes A, M \otimes A') \hto (A, A')\] which is given by $\langle \id_{M \otimes A}\mid \id_{M \otimes A'} \rangle$.
\end{example}

\begin{remark}
  Elements of $\Optic_\C((S, S'), (A, A'))$ have an appealing
  interpretation as string diagrams with a ``hole'' missing. We draw the
  pair $\langle l \mid r \rangle$ as
  \begin{center}
    \input{diagrams/generic-optic.tikz}
  \end{center}
  reading left to right, so the left half of the diagram represents $l$ and the right half $r$. The relation imposed by the coend equates the following two diagrams:
  \begin{center}
    \input{diagrams/coend-relation-left.tikz}
    \qquad
    \input{diagrams/coend-relation-right.tikz}
  \end{center}
  We will therefore omit the vertical cut point between $l$ and $r$ in subsequent diagrams; any choice yields a representative of same optic.
\end{remark}

Two optics can be composed as follows: \todo{I give possibly too much detail here.} We wish to construct a map
\begin{align*}
  &\left(\int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')\right) \times \left(\int^{M \in \C} \C(R, M \otimes S) \times \C(M \otimes S', R')\right) \\ &
                                                                                                                                                                       \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
The product in $\Set$ preserves colimits, so in particular coends. Using this fact, and the Fubini theorem for coends~\cite[Remark 1.9]{CoendCofriend}, the domain is isomorphic to
\begin{align*}
  \int^{(M, N) \in \C \times \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R').
\end{align*}
So by the universal property of coends, it suffices to construct natural maps
\begin{align*}
  & \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') \\ &
                                                                                                              \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
for each $M, N \in \C$. For this we use the composites
\begin{align*}
  &\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R')\\
  \to \,& \C(N \otimes S, N \otimes M \otimes A) \times \C(N \otimes M \otimes A', N \otimes S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') && \text{(functoriality of $N \otimes  -$)} \\
  \to \,& \C(R, N \otimes  M \otimes A) \times \C(N \otimes M \otimes A', R') && \text{(composition in $\C$)} \\
  \to \,&\int^{M \in \M} \C(R, M \otimes A) \times \C(M \otimes A', R') && \text{($\copr_{N \otimes M}$)}
\end{align*}
In other words, let $\langle l', r' \rangle : (R, R') \hto (S, S')$ and $\langle l, r \rangle : (S, S') \hto (A, A')$ with $M$ the complement for $\langle l', r' \rangle$. The composite is then: $\langle l, r \rangle \circ \langle l', r' \rangle = \langle (M l)l', r'(Mr)\rangle$.

@@

\begin{proposition}
  The above data form a category $\Optic_\C$.
\end{proposition}
\begin{proof}
  In~\cite[Section 6]{Doubles} this is proven abstractly, by exhibiting this category as the Kleisli category for a monad in the bicategory $\Prof$. Here we prove the properties directly.

  We check associativity of composition equationally. As in the definition of composition, the Fubini theorem for coends allows us to choose representatives for all three optics simultaneously. Suppose we have representatives of three optics $\langle l_1, r_1\rangle : (R, R) \hto (S, S')$ and $\langle l_2, r_2 \rangle : (S, S') \hto (A, A')$ and $\langle l_3, r_3 \rangle : (A, A') \hto (B, B')$, that have complements $M$, $N$ and $P$ respectively. Then:
  \begin{align*}
    (\langle l_3, r_3 \rangle \circ \langle l_2, r_2 \rangle) \circ \langle l_1, r_1 \rangle
    &= \langle (N \otimes l_3)l_2, r_2(N \otimes r_3) \rangle \circ \langle l_1, r_1 \rangle \\
    &= \langle (M \otimes ((N \otimes l_3)l_2))l_1, r_1(M \otimes (r_2(N \otimes r_3))) \rangle \\
    &= \langle (M \otimes N \otimes l_3)(M \otimes l_2)l_1, r_1(M \otimes r_2)(M \otimes N \otimes r_3) \rangle \\
    &= \langle l_3, r_3 \rangle \circ (\langle (M \otimes l_2)l_1, r_1(M \otimes r_2) \rangle) \\
    &= \langle l_3, r_3 \rangle \circ (\langle l_2, r_2 \rangle \circ \langle l_1, r_1 \rangle)
  \end{align*}

@@

The remainder of this section comprises some useful observations that were not made in~\cite{Doubles}.

@@

\begin{remark}
  This functor is not necessarily faithful, see Remark~\ref{lens-iota-not-faithful}.
\end{remark}

% \todo{This doesn't give us the structure of a `framing', we can only lift vertical arrows to horizontal arrows facing one direction.}

% \begin{theorem}
%   $\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, S' \otimes T')$, the unit is $(I, I)$, and the action on morphisms induced by
%   \begin{align*}
      %       &\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
      %       \to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
      %       \to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
      %       \to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}
                                                                                                                                    %     \end{align*}
                                                                                                                                    %                                                                                                                                     \end{theorem}
                                                                                                                                    %                                                                                                                                     \begin{proof}
                                                                                                                                    %                                                                                                                                     Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
                                                                                                                                    %                                                                                                                                     \begin{align*}
                                                                                                                                    %                                                                                                                                     \langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
                                                                                                                                                                                                                                                                                                                                %     \end{align*}
                                                                                                                                                                                                                                                                                                                                %                                                                                                                                                                                                                                                                                                                                 This is functorial, as suggested by the equivalence of the following string diagrams:
                                                                                                                                                                                                                                                                                                                                %                                                                                                                                                                                                                                                                                                                                 \todo{todo}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 For future convenience, note that
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \begin{align*}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   \langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &= \langle l \otimes T, r \otimes T' \rangle
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \end{align*}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 Now we check functoriality. Suppose we have:
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \begin{align*}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   \langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   \langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   \langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   \langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \end{align*}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \begin{align*}
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   &\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                   = \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
                                                                                                                                                                                                                                                                                                                                %%                                                                                                                                                                                                                                                                                                                                 \end{align*}

@@

                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            %                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             The structure morphisms are all lifted from the structure morphisms in $\C$:
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            %                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             \begin{align*}
                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                            %                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                             \alpha_{(R, R'), (S, S'), (T, T')} &:= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1}) \\
      %       \lambda_{(S, S')} &:= \iota(\lambda_{(S, S')}, \lambda_{(S, S')}^{-1}) \\
      %       \rho_{(S, S')} &:= (\rho_{(S, S')}, \rho_{(S, S')}^{-1}) \\
      %       s_{(S, S'), (T, T')} &:= \iota(s_{S, T}, s_{T', S'})
                                     %     \end{align*}

@@

                                     %                                      Note that because $\iota(S, S') = (S, S')$, required equations for $\iota$ to be a monoidal functor hold by definition (although we don't yet know that $\Optic_\otimes$ is monoidal). The pentagon and triangle axioms then hold in $\Optic_\otimes$, as they are the image of the same axioms in $\C \times \C^\op$ under $\iota$. The only remaining thing to verify is that these structure maps are natural in $\Optic_\otimes$.

@@

                                     %                                      Let us check that $\alpha_{(R, R'), (S, S'), (T, T')}$ is natural in $(R, R')$. Suppose $f : (Q, Q') \hto (R, R')$ has representative $(l, r) : (Q \to M \otimes R, M \otimes R' \to Q')$, then:
                                     %                                      \begin{align*}
                                     %                        &\alpha_{(R, R'), (S, S'), (T, T')} \circ (f \otimes (S, S')) \otimes (T, T') \\
      %               &= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1})  \circ (\langle l, r \rangle \otimes (S, S')) \otimes (T, T') \\
      %               &= \langle\lambda_{R \otimes (S \otimes T)}\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1} \lambda^{-1}_{(R' \otimes S') \otimes T'} \rangle  \circ \left\langle (l \otimes S) \otimes T, (r \otimes S') \otimes T' \right\rangle \\
